Quantum H theorem, variational entropy, and relaxation processes

### 2020-12-18 tex/classical.tex
Rewording: start of section 2.3

We emphasize that the validity of the $H$-theorem relies significantly on assuming
that the distribution function is homogeneous and that the hypothesis of
molecular chaos is surmised.

@@

The Boltzmann $H$-functional (\ref{eq:hfunctional}) is defined with a spatially
homogeneous distribution $f(\vec{v},t)$. However, we want to consider a $H$
functional with a non-homogeneous distribution function in order to generalize
the Boltzmann $H$-theorem. Because of this, we introduce a modified $H$
functional (denoted as $H'$) defined as
%
\begin{equation}\label{eq:CH2}
   H'(t)=\sum_{M=1}^{K}\int f_M(\vec{v},t) \ln f_M(\vec{v},t)\mathrm{d}^3v.
\end{equation}
%
This $H'$ describes a system with a global energy $E$, a global number of free
classical particles $N$, global volume $V$, and a temperature $T$. Nevertheless,
the global system, in turn, consists of a set of $K$ small cells with a constant
volume $\delta V_M$, whose local homogeneous distribution functions is
$f_{M}(\vec{v},t)$.

The distribution function will depend on the position of the cells, the velocity
$\vec{v}$, and time $t$. Without loss of generality, we choose these cells to
have local identical volumes (\textit{i.e.} $\delta V_M = V/K, \forall M$).
Each cell has its local chemical potential, $\mu_M$, local
$H$ functional, $H'_M$, local number of particles, $\mathcal{N}'_M$, and local
energy, $\mathcal{E}'_M$. In terms of the local distribution 
function and the energy $\epsilon(\vec{v})$, the last three properties are
given by:
%
\begin{eqnarray}
    H_M' & = &  \int f_M(\vec{v},t) \ln f_{M}(\vec{v},t)
      \mathrm{d}^{3}v \label{Hcell},\nonumber\\
    \mathcal{N}_M' & = & \int f_{M}(\vec{v} ,t) \mathrm{d}^{3}v, \nonumber\\
    \mathcal{E}_M' & = & \int f_{M}(\vec{v},t)\epsilon(\vec{v}) \mathrm{d}^{3}v.
\end{eqnarray}
%

When the system is in equilibrium, the local number of particles and the local
energy do not depend on the cell number, this is
%
\begin{equation}
   \mathcal{N}'_M=\mathcal{N}'\equiv \bar{\mathcal{N}}';
   \quad\mathcal{E}'_M=\mathcal{E}'\equiv\bar{\mathcal{E}}'.
\end{equation}
%
The particles in the system are considered to be free, \textit{i.e.} the
particles do not interact with each other. Besides, the local distribution
function must satisfy the following microcanonical restrictions 
%
\begin{equation}\label{micro}
    \sum_{M=1}^{K}\int f_M(\vec{v},t)\mathrm{d}^3v =N,
    \quad\sum_{M=1}^{K}\int f_M(\vec{v},t)\epsilon(\vec{v})\mathrm{d}^3v=E.
\end{equation}
%

>>>>

@@

As we saw in the previous section,
the validity of the $H$-theorem relies significantly on assuming
that the distribution function is homogeneous and that the hypothesis of
molecular chaos is surmised. However, in order to consider systems whose
distribution functions are not necessarily homogeneous, which
allows for studying systems out of equilibrium,
we introduce a modified $H$ functional as follows.

The proposed functional, denoted as $H'$, describes a diluted gas occupying a volume, $V$,
divided into $K$ cells, which, without loss of generality,
have local identical volumes (\textit{i.e.} $\delta V_M = V/K,\ M=1,\,\dots\,,K$).
Each cell has its own local
$H$ functional, $H'_M$, and local homogeneous distribution function,
$f_{M}(\vec{v},t)$; local number of particles, $\mathcal{N}'_M$;
local temperature, $T_M$; and a local energy, $\mathcal{E}'_M$. As a whole, the system has a
global energy $E$, and a global number of free
classical particles $N$. The functional takes the form
%
\begin{equation}\label{eq:CH2}
   H'(t)=\sum_{M=1}^{K}\int f_M(\vec{v},t) \ln f_M(\vec{v},t)\mathrm{d}^3v.
\end{equation}
%

The distribution functions, $\{f_M(\vec{v},t)\}$, depend implicitly on the position of the cells,
relative to the global system, and on the velocity
$\vec{v}$ and time $t$. In terms of $f_M(\vec{v},t)$ and the energy, $\epsilon(\vec{v})$, we have:
%
\begin{subequations}
\begin{eqnarray}
    H_M' & = &  \int f_M(\vec{v},t) \ln f_{M}(\vec{v},t)
      \mathrm{d}^{3}v \label{Hcell},\\
    \mathcal{N}_M' & = & \int f_{M}(\vec{v} ,t) \mathrm{d}^{3}v,\\
    \mathcal{E}_M' & = & \int f_{M}(\vec{v},t)\epsilon(\vec{v}) \mathrm{d}^{3}v.
\end{eqnarray}
\end{subequations}
%
Notice that assuming local homogeneous distribution functions means that we are surmising the 
local equilibrium hypothesis. In addition, $f_{M}(\vec{v},t)$ must satisfy the following
microcanonical restrictions:
%
\begin{subequations}\label{eq:micro}
\begin{equation}\label{eq:micron}
    \sum_{M=1}^{K}\int f_M(\vec{v},t)\mathrm{d}^3v =N
\end{equation}
and
\begin{equation}\label{eq:microe}
    \sum_{M=1}^{K}\int f_M(\vec{v},t)\epsilon(\vec{v})\mathrm{d}^3v=E.
\end{equation}
\end{subequations}
%

If the complete system is in equilibrium, then the local number
of particles and the local
energy do not depend on the cell number, this is
%
\begin{subequations}\label{eq:eqNandE}
\begin{equation}
   \mathcal{N}'_M=\mathcal{N}'\equiv \bar{\mathcal{N}}'
\end{equation}
%
and
%
\begin{equation}
   \quad\mathcal{E}'_M=\mathcal{E}'\equiv\bar{\mathcal{E}}'.
\end{equation}
\end{subequations}
%
In Eqs. (\ref{eq:eqNandE}) the bar implies averaged properties over the complete system.
